---
title: "TENR Zebra SCR analysis - search path analysis"
output: html_notebook
date: 2022-06-21
---

## setup workspace
```{r load libraries}
library(tidyverse)
library(lubridate)
library(sp)
library(rgdal)
library(rgeos)
library(scrbook)
library(lattice)
library(coda)
library(nimble)
library(mcmcplots)
library(abind)
library(here)
```

## load data, project in UTM-km
```{r load data}
# coordinate systems
crs.dd <- CRS('+init=epsg:4326') # WGS84 DD
crs.utm <- CRS('+init=epsg:32735') # WGS84 UTM 35S
crs.utmkm <- CRS('+init=epsg:32735 +units=km') # WGS84 UTM 35S units = km

# data set
zebEncs.raw <- read_csv(here('Data', 'ZebEncsCombined.csv'))
zebEncs.red <- zebEncs.raw %>%
  select(encID = EncID, date = Date, time = Time, ageCat = AgeCat,
         indID = IndID, dist = Distance, xcoord = Xzeb, ycoord = Yzeb) %>%
  mutate(occID = case_when(date == '2017-07-31' ~ 1,
                           date == '2017-08-01' ~ 2,
                           date == '2017-08-02' ~ 3,
                           date == '2017-08-03' ~ 4,
                           date == '2017-08-04' ~ 5,
                           date == '2017-08-06' ~ 6,
                           date == '2017-08-07' ~ 7,
                           date == '2017-08-08' ~ 8,
                           date == '2017-08-09' ~ 9,
                           date == '2017-08-10' ~ 10)) %>%
  distinct(indID, occID, .keep_all= TRUE) %>%
  mutate(indID = factor(indID)) %>%
  mutate(occID = factor(occID)) %>%
  arrange(indID, occID)
coordinates(zebEncs.red) <- c('xcoord', 'ycoord')
proj4string(zebEncs.red) <- crs.dd
zebEncs.utmkm <- spTransform(zebEncs.red, crs.utmkm)
zebEncs.df <- zebEncs.utmkm@data
```

## state-space, project in UTM-km
```{r state-space}
SS <- readOGR('GIS', layer = 'TENRStateSpace')
SS <- spTransform(SS, crs.utmkm)

Xl <- bbox(SS)[1,1]
Xu <- bbox(SS)[1,2]
Yl <- bbox(SS)[2,1]
Yu <- bbox(SS)[2,2]

# Checking
#plot(SS, axes = T)
#plot(zebEncs.utmkm, add = T)
```

## load and organise data for analysis

y: observations, 0s and 1s, i(ndividuals) * k(occasions) matrix

T: no. occasions (= K = nocc)
```{r y: obs; T: no. occasions}
y <- table(zebEncs.df$indID, zebEncs.df$occID)
y <- matrix(y, ncol = ncol(y))
ysum <- apply(y, 1, sum)
y <- y[ysum > 0,]
nocc <- dim(y)[2]
```

u: observed x-coordinates + NAs, i * k matrix
```{r u: observed x-s}
zeb.u <- cbind(zebEncs.utmkm@data, xcoord = zebEncs.utmkm@coords[,1]) %>%
  select(indID, occID, xcoord) %>%
  spread(key = occID, value = xcoord)
zeb.u <- as.matrix(zeb.u[,2:ncol(zeb.u)])
```

v: observed y-coordinates + NAs, i * k matrix
```{r v: observed y-s}
zeb.v <- cbind(zebEncs.utmkm@data, ycoord = zebEncs.utmkm@coords[,2]) %>%
  select(indID, occID, ycoord) %>%
  spread(key = occID, value = ycoord)
zeb.v <- as.matrix(zeb.v[,2:ncol(zeb.v)])
```

X: coordinates of the search path (ends of each segment)

J: no. segments (= trap locations)
```{r X: search paths; J: no. segments}
# one path per occasion; load path, thin by designated amount
# median path length: ~63 km
# 42 pts is 1 point every ~1.50 km
# 50 pts is 1 point every ~1.25 km
# 63 pts is 1 point every ~1.00 km
# 84 pts is 1 point every ~0.75 km
# 126 pts is 1 point every ~0.50 km -- errors out on the cluster
# Chosen to match 1-km grid on grid-based analysis; not sure how important this is
# Smaller than 2-sigma = ~1.6 km from prelim analysis
# Smaller resolution than this produces memory and processing problems

J <- 42
pathFiles <- list.files(path = 'GIS', pattern = 'occ') %>%
  str_sub(1, 9) %>%
  unique()

processPaths <- function(path, pathFiles, J, K, coordsys){
  X <- array(NA, dim = c(J, 2, K)) # empty array to hold coordinates
  for(i in 1:length(pathFiles)){
    pathLayer <- readOGR(dsn = path, layer = pathFiles[i])
    pathLayer <- spTransform(pathLayer, coordsys)
    segs <- spsample(pathLayer, n = J, type = 'regular')
    X[,,i] <- coordinates(segs)
  }
  return(X)
}

X <- processPaths(path = 'GIS', pathFiles = pathFiles, J = J, K = nocc,
                  coordsys = crs.utmkm)

# For a figure
#occ10Path <- readOGR('GIS', layer = 'occ10Path')
#occ10Path <- spTransform(occ10Path, crs.utmkm)
#plot(SS, lwd = 1, axes = T, xlab = 'km E', ylab = 'km N')
#plot(occ10Path, lty = 2, add = T)
```

M: data augmentation
```{r M: data augmentation}
M <- 1500
nz <- M - nrow(y)
yAug <- rbind(y, matrix(0, nrow = nz, ncol = ncol(y)))
u <- rbind(zeb.u, matrix(NA, nrow = nz, ncol = ncol(y)))
v <- rbind(zeb.v, matrix(NA, nrow = nz, ncol = ncol(y)))

sx <- runif(M, Xl, Xu)
sy <- runif(M, Yl, Yu)
S <- cbind(sx, sy)
```

## setup for BUGS
```{r setup for BUGS}
# rescale data
X.rsc <- array(NA, dim = c(J, 2, nocc))
X.rsc <- abind(X[,1,] - bbox(SS)[1,1], X[,2,] - bbox(SS)[2,1], along = 3)
X.rsc <- aperm(X.rsc, c(1, 3, 2))
S.rsc <- matrix(NA, nrow = nrow(S), ncol = 2)
S.rsc <- cbind(S[,1] - bbox(SS)[1,1], S[,2] - bbox(SS)[2,1])

# data, with rescaling
my.data <- list(y = yAug,
                u = u - bbox(SS)[1,1],
                v = v - bbox(SS)[2,1],
                X = X.rsc)

my.const <- list(Xl = Xl - bbox(SS)[1,1],
                 Xu = Xu - bbox(SS)[1,1],
                 Yl = Yl - bbox(SS)[2,1],
                 Yu = Yu - bbox(SS)[2,1],
                 K = nocc, M = M, J = J)

# initial values
s.init <- S.rsc
w.init <- c(rep(1, (M - nz)), rbinom(nz, 1, 0.5))

my.inits <- list(beta0 = runif(1, -1, 5),
                 beta1 = runif(1, 4, 10),
                 lsigma = runif(1, -1, 0.5),
                 s = s.init, w = w.init)

# parameters to be estimated
#my.params <- c('beta1', 'N', 'psi', 'sigma', 'beta0',
#            'X1obs', 'X1new', 'X2obs', 'X2new')
my.params <- c('beta0', 'beta1', 'N', 'psi', 'sigma', 's', 'w', 'T1obs', 'T1new')

zebData <- list(my.data, my.const, s.init, w.init, my.params, X.rsc)
save(zebData, file = 'spacing150.RData', version = 2)
```
